---
title: "Gender Parity of TIME Magazine covers using plotly"
author: "Ian Roberts"
date: "April 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
```

##Background
What is being visualized in this project is the Gender breakdown of TIME Magazine covers by crowdflower found [here](https://data.world/crowdflower/time-magazine-covers). From the description: 'Contributors were shown images of Time Magazine covers since the late 1920s and asked to classify if the person was male or female. Data is broken down overall and on an annual basis.'

## By Total Number of Apperances
```{r thing}
library(dplyr)
library(tidyr)
library(plotly)

tcv <- read.csv("data/TIME_Gender_Ratio.csv", stringsAsFactors = F)
long <- tcv %>% 
        select(Male, Female) %>%
        gather %>%
        mutate(year = rep(ts(tcv$Year), 2)) %>%
        mutate(total = rep(tcv$Total, 2))

plot_ly(x = long[,3], y = long[,2], color = long[,1], type = 'scatter') %>%
        add_trace(y = long[,4], name = 'Total', type = 'scatter', mode = 'lines', line = list(color = 'rgb(205,12,24)', dash = 'dash')) %>%
                layout(title = 'Gender of Person on TIME Magazine Cover',
                        yaxis = list(title = "Total Yearly Apperances"))
```


## By Percentage of Appearances
```{r thing2}
long <- tcv %>% 
        select(Male.., Female..) %>%
        gather

long$value <- gsub("%", "", long$value) %>% as.numeric

long <- long %>% 
        mutate(year = rep(ts(tcv$Year), 2))

longF <- long[long$key == 'Female..',]
longM <- long[long$key == 'Male..',]


plot_ly(x = longF$year, y = longF$value, type = 'scatter', mode = 'none', name = 'Female', fill = 'tozeroy') %>%
        add_trace(x = longM$year, y = longM$value, name = 'Male', mode = 'none', fill = 'tonexty') %>% 
                layout(title = 'Gender of Person on TIME Magazine Cover',
                yaxis = list(title = "% of Yearly Apperances"))

```

##Conclusion

To quote [the article](https://www.crowdflower.com/time-magazine-cover-data/) in which this dataset is referenced: "*TIME* magazine really likes dudes"